### 2017-01-02 paper/paris17/paper.tex
Weijia revise it

%As the Earth attenuates the seismic signals in real Earth materials, an accuracy wavefield modeling approach needs to account for the effects of attenuation and velocity dispersion \cite[]{zhu2014theory}.

@@

The concept of the constant-Q model introduced by \cite{Kjartansson.sep.23} is used to approximate the attenuation. However, the formulation for computing the time derivative of the constant-Q model requires to store wavefields from previous values to present time \cite[]{podlubny1998fractional}, which is impractical for real seismic modeling in terms of memory capacities on modern computers. Zhu developed a constant-Q wave equation using fractional Laplacian operators which avoid the need for storing the past values of the wavefields \cite[]{zhu2014theory}. However, the proposed method is for 2D viscoelastic case and the computational cost of fractional Laplacian is still expensive compared with the conventional Laplacian.

Meanwhile, the Nvidia Graphic Processing Units (GPUS)  have achieved a great success in many domains. In seismic applications, GPUs are employed as high-performance computing accelerators for key seismic modeling, imaging and inversion kernels due to their high memory bandwidth and a large number of processing units \cite[]{he2015gpu}.

@@

We extend the formulation of \cite{zhu2014theory} from the 2D viscoelastic case to the 3D viscoelastic case. Note that the Laplacian differential operators of fractional order bring extra computational costs to the elastic modeling, we approximate it with the conventional Laplacian operators, which significantly reduces the computing complexities. Also, we further reduce the computational costs by ignoring the dispersion terms because it makes little effects compared to the attenuation of the signals. As a result, our equations are as follows.

@@

where $\sigma$ is the stress tensor; $\epsilon$ is the strain tensor; $c_{ij}$ is the forth-order stiffness tensor; $C_p$ and $C_s$ are the velocities of P-wave and S-wave. $Q_{p,s}$ is the constant factor for the attenuation of P-wave and S-wave respectively. Note that we only present how $\tau_p^{(1)}$ and $\tau_s^{(1)}$ are derived, $\tau_p^{(2)}$,$\tau_p^{(3)}$, $\tau_s^{(2)}$ and $\tau_s^{(3)}$ can be derived in a similar way.

@@

The reason why GPUs outperforms CPUs significantly on many scenarios is that the architecture of GPUs favors the throughput of many data-parallel tasks over the latency of a single thread. We have an efficient parallel design on the multiple-GPU platform with three main strategies.

Firstly, we fully utilize the hierarchical memory of the GPU. We keep a set of constant variables such as the derivative coefficients in the constant memory and the read-only cache. As the 64KB first-level cache can be configured as different combinations between L1 cache and shared memory, we configure it dynamically in different scenarios. For the 8th-order stencil computation, we configure the first-level cache by setting the sizes of shared memory and L1 cache to be 48 KB and 16 KB respectively, which achieves great performance based on the 3D stencil design proposed by \cite{micikevicius20093d}. For others, we prefer to configure the first-level cache as the L1 cache where we don't use shared memory explicitly.

@@

Thirdly, we design a computation and communication overlapping scheme to maximize the utilization of different kinds of resources. The entire wavefield domain is divided into 4 subdomains, each assigning to a GPU shown in Figure \ref{fig:domain-decomposition}. Note that we need ghost cells to store and exchange the halos, denoted by the arrows. In each subdomain, we further divide it into three parts, the halo part, the outer part and the inner part as shown in Figure \ref{fig:task-decomposition}. The computation of inner and outer parts are assigned to different CUDA streams so that they are executed concurrently. When the computation of the outer part is finished, the GPU starts exchanging halos with its neighbors. In this way, the computation of the inner part can hide the communication time perfectly, shown in Figure \ref{fig:overlap}.

@@

\subsection{Q Effects}
Experiments for demonstrating the functionality of applying the Q effect and the performance with different optimization strategies are presented. Figure \ref{fig:with-attenuation} shows the z-x plane of the z-component of the 3D wavefield using standard elastic wave equation at time 0.9 second while the Figure \ref{fig:without-attenuation} shows the wavefields with the constant Q ($Q_s=Q_p=20$) added at the same time step. We can see that the wavefields are nearly identical except that the wavefront in Figure \ref{fig:with-attenuation} is weaker, which is reasonable owning to the energy attenuation. The spectrums of the wavefields give a better illustration for the attenuation, which is shown in Figure \ref{fig:spectrum}.

@@

We propagate the wavefield in the 3D domain of size $700\times700\times700$ for 6000 time steps to measure the performance. Figure \ref{fig:gpu-speedup} shows the performance speedup of 1-4 K40 GPUs over one 24-core CPU node. The GPU design among 4 K40 GPUs can significantly improve the performance of the computational part by 96 times, and improve the overall performance by 63 times. Note that the CPU design is also well tuned by using multi-threading and vectorization. Figure \ref{fig:q-speedup} shows the speedup as a function of propagation time for different Q values. We can see that we obtain a great performance gain at the early and late time steps because we restrict the computing region at the early time steps and ignore the high frequencies at the late time steps. The longer the time records, the more significant speedup we achieve.  Combing all optimization schemes as well as the GPU-based optimizations, we can achieve a performance speedup of 60x to 200x as a function of Q.

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.4\textwidth}
        \centering
        \includegraphics[height=1.5in]{./fig/speedup.eps}
        \caption{Speedup with GPU optimizations}
        \label{fig:gpu-speedup}
    \end{subfigure}%
    ~
    \begin{subfigure}[b]{0.4\textwidth}
        \centering
        \includegraphics[height=1.5in]{./fig/speedup_q.eps}
        \caption{Speedup as different Q value}
        \label{fig:q-speedup}
    \end{subfigure}
    \caption{Performance result of GPU optimizations and energy attenuation}
\end{figure}

@@

This work manages to boost the performance the wavefield propagation in 3D elastic scenarios by approximating the Q propagation and the using multi-GPU platform. We first extend the constant-Q formulation from the 2D viscoelastic to the 3D viscoelastic case. Different optimization techniques on GPUs are then described for an efficient modeling kernel. We also propose a set of schemes to reduce the computation to further improve performance. The experimental results show that we can achieve significant performance speedup of 60 to 200 times with 4 GPUs over the CPU-based solution as a function of Q.

>>>>

@@

The concept of the constant-Q model introduced by \cite{Kjartansson.sep.23} is used to approximate the attenuation. However, the formulation for computing the time derivative of the constant-Q model requires to store wavefields from previous values to present time \cite[]{podlubny1998fractional}, which is impractical for real seismic modeling in terms of memory capacities on modern computers. Zhu developed a constant-Q wave equation using fractional Laplacian operators which avoid the need for storing the past values of the wavefields \cite[]{zhu2014theory}. However, the proposed method is for the 2D viscoelastic case and the computational cost of fractional Laplacian is still expensive compared with the conventional Laplacian.

Meanwhile, the Nvidia Graphic Processing Units (GPUs)  have achieved a great success in many domains. In seismic applications, GPUs are employed as high-performance computing accelerators for key seismic modeling, imaging and inversion kernels due to their high memory bandwidth and a large number of processing units \cite[]{he2015gpu}.

@@

We extend the formulation of \cite{zhu2014theory} from the 2D viscoelastic case to the 3D viscoelastic case. Note that the Laplacian differential operators of fractional order bring extra computational costs to the elastic modeling, we approximate it with the conventional Laplacian operators, which significantly reduces the computing complexities. Also, we further reduce the computational costs by ignoring the dispersion terms because it makes little effects compared with the attenuation of the signals. As a result, our equations are as follows.

@@

where $\sigma$ is the stress tensor; $\epsilon$ is the strain tensor; $c_{ij}$ is the forth-order stiffness tensor; $C_p$ and $C_s$ are the velocities of P-wave and S-wave. $Q_{p,s}$ is the constant factor for the attenuation of P-wave and S-wave, respectively. Note that we only present how $\tau_p^{(1)}$ and $\tau_s^{(1)}$ are derived$. \tau_p^{(2)}$,$\tau_p^{(3)}$, $\tau_s^{(2)}$ and $\tau_s^{(3)}$ can be derived in a similar way.

@@

The reason why GPUs outperform CPUs significantly on many scenarios is that the architecture of GPUs favors the throughput of many data-parallel tasks over the latency of a single thread. We have an efficient parallel design on the multiple-GPU platform with three main strategies.

Firstly, we fully utilize the hierarchical memory of the GPU. We keep a set of constant variables such as the derivative coefficients in the constant memory and the read-only cache. As the 64KB first-level cache can be configured as different combinations between L1 cache and shared memory, we configure it dynamically in different scenarios. For the 8th-order stencil computation, we configure the first-level cache by setting the sizes of shared memory and L1 cache to 48 KB and 16 KB, respectively, which achieves great performance based on the 3D stencil design proposed by \cite{micikevicius20093d}. For others, we prefer to configure the first-level cache as the L1 cache where we don't use shared memory explicitly.

@@

Thirdly, we design a computation and communication overlapping scheme to maximize the utilization of different kinds of resources. The entire wavefield domain is divided into 4 subdomains, each assigning to a GPU shown in Figure \ref{fig:domain-decomposition}. Note that we need ghost cells to store and exchange the halos, which is denoted by the arrows. In each subdomain, we further divide it into three parts, the halo part, the outer part and the inner part as shown in Figure \ref{fig:task-decomposition}. The computation of inner and outer parts is assigned to different CUDA streams so that they are executed concurrently. When the computation of the outer part is finished, the GPU starts exchanging halos with its neighbors. In this way, the computation of the inner part can hide the communication time perfectly, shown in Figure \ref{fig:overlap}.

@@

\subsection{Q Effects}
Experiments for demonstrating the functionality of applying the Q effect and the performance with different optimization strategies are presented. Figure \ref{fig:with-attenuation} shows the z-x plane of the z-component of the 3D wavefield using standard elastic wave equation at 0.9 seconds while Figure \ref{fig:without-attenuation} shows the wavefields with the constant Q ($Q_s=Q_p=20$) added at the same time step. We can observe that the wavefields are nearly identical except that the wavefront in Figure \ref{fig:with-attenuation} is weaker, which is reasonable owning to the energy attenuation. The spectrums of the wavefields give a better illustration for the attenuation, which is shown in Figure \ref{fig:spectrum}.

@@

We propagate the wavefield in the 3D domain of size $700\times700\times700$ for 6000 time steps to measure the performance. Figure \ref{fig:gpu-speedup} shows the performance speedup of 1-4 K40 GPUs over one 24-core CPU node. The GPU design among 4 K40 GPUs can significantly improve the performance of the computational part by 96 times, and improve the overall performance by 63 times. Note that the CPU design is also well tuned by using multi-threading and vectorization. Figure \ref{fig:q-speedup} shows the speedup as a function of propagation time for different Q values. We can see that we obtain a great performance gain at the early and late time steps because we restrict the computing region at the early time steps and ignore the high frequencies at the late time steps. The longer the time records, the more significant speedup we achieve. Combining all optimization schemes as well as the GPU-based optimizations, we can achieve a performance speedup of 60x to 200x as a function of Q.

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.4\textwidth}
        \centering
        \includegraphics[height=1.5in]{./fig/speedup.eps}
        \caption{Speedup with GPU optimizations}
        \label{fig:gpu-speedup}
    \end{subfigure}%
    ~
    \begin{subfigure}[b]{0.4\textwidth}
        \centering
        \includegraphics[height=1.5in]{./fig/speedup_q.eps}
        \caption{Speedup as different Q value}
        \label{fig:q-speedup}
    \end{subfigure}
    \caption{Performance results of GPU optimizations and energy attenuation}
\end{figure}

@@

This work manages to boost the performance of the wavefield propagation in 3D elastic scenarios by approximating the Q propagation and using the multi-GPU platform. We first extend the constant-Q formulation from the 2D viscoelastic case to the 3D viscoelastic case. Different optimization techniques on GPUs are then described for an efficient modeling kernel. We also propose a set of schemes to reduce the computation to further improve the performance. The experimental results show that we can achieve significant performance speedups of 60 to 200 times with 4 GPUs over the CPU-based solution as a function of Q.
